---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# PipMet

<!-- badges: start -->

The PipMet package was developed to process begin-to-end metabolomic-based GC-MS data, with the automatized generation of pictures throughout the processing in high quality. Every input from the user is taken through pop-up windows.This package was firstly developed for the Brazilian National Biorenewables Laboratory (LNBR) from the Brazilian National Center for Research in Energy and Materials (CNPEM).

<!-- badges: end -->

## Installation

You can install the released version of PipMet from [GitHub](https://github.com) with:

``` r
devtools::install_github("AnnafCouto/PipMet")
```

## Example

The package is constituted of one main function with pre-set parameters and algorithms. To start the processing:

```{r example}
library(PipMet)
```

The package was thought to be the more friendly-user as possible, even though it runs in code lines. Therefore, when information is needed, pop-ups will appear so that the user can input them.

```{r example}
GC_dataProcess(
   sample_dir = system.file("extdata", package = "PipMet"),
   metadata = system.file("extdata", "metadata.csv", package = "PipMet"),
   extensao = ".mzML",
   myDir = '~/',
   example = TRUE,
   pictures = FALSE
)
```
Set 'pictures = TRUE' to generate pictures along the code.

For more information, see the package [vignette](vignettes/workData.Rmd).
