---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# PipMet

<!-- badges: start -->

The PipMet package was developed to process begin-to-end metabolomic-based GC-MS data, with the automatized generation of pictures throughout the processing in high quality. Every input from the user is taken through pop-up windows.This package was firstly developed for the Brazilian National Biorenewables Laboratory (LNBR) from the Brazilian National Center for Research in Energy and Materials (CNPEM).

<!-- badges: end -->

## Installation

You can install the released version of PipMet from [GitHub](https://github.com) with:

``` r
devtools::install_github("AnnafCouto/PipMet")
```

## Example

The package is constituted of one main function with pre-set parameters and algorithms. To start the processing:

```{r example}
library(PipMet)
{GC_dataProcess()}
```

