---
title: "GC-MS Data Processing wiht PipMet"
author: "Anna Couto"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GC-MS Data Processing wiht PipMet}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction
---

This vignette describe the usage of the global function `workData()` from the PipMet R package in order to process GC-MS acquired data. 
The `workData()` is a wrapper for read files, preprocessing the raw data into viable spectra and a quantification table and normalizing the results. The user may provide the informations required through parameters to the `workData()` function or provide them through pop-up windows. 

Therefore, the only function the user needs to call is as follows:

```{r, echo =TRUE, results = 'hide', eval = FALSE}
library(PipMet)
result <- workData()
```

Information regarding the processing may be passed through parameters to the `workData()` function, according to the documentation, so that the pop-ups don't show.

## Data processing example

The PipMet R package accompanies four GC-MS analysis of sugarcane colms and its commercial variety energycane for examples. The samples were collected and analysed as described by Abreu and colleges^1^.



## Files reading
---

The PipMet package is capable of reading '.mzML' and '.mzXML' files through the xcms, MSnbase and mzR packages. 

## Figures
---

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))

## References

1. Abreu LGF, Silva N, Ferrari A, Carvalho L, Fiamenghi M, Carazolle M, Fill T, Pilau E, Pereira G, Grassi M: **Metabolite profiles of energy cane and sugarcane reveal different strategies during the axillary bud outgrowth**. *Plant Physiology and Biochemistry*, 2021, ***167***:504-516. 
